\section{Notation}
\label{app:notation}

\renewcommand{\arraystretch}{1.12}
\begin{longtable}{@{}p{0.25\textwidth}p{0.68\textwidth}@{}}
\toprule
\textbf{Notation} & \textbf{Meaning} \\
\midrule
\endfirsthead

\toprule
\textbf{Notation} & \textbf{Meaning} \\
\midrule
\endhead

$d$ & Ambient dimension of the primal variable $x\in\RR^d_+$. \\
$A=(A_1;A_2)$ & Constraint matrix split into two blocks. \\
$b=(b_1;b_2)$ & Right-hand side split compatibly with $A_1,A_2$. \\
$C\in\RR^d$ & Linear cost vector in the unregularized linear program. \\
$\z\in\RR^d_{++}$ & Positive reference measure/vector for the KL penalty. \\
$\gamma>0$ & Entropic regularization parameter. \\
$\Cc_1,\Cc_2$ & Affine constraint blocks $\{A_1x=b_1\}$ and $\{A_2x=b_2\}$. \\
$\KLdiv{x}{z}$ & Non-normalized Kullback--Leibler divergence from $x$ to $z$. \\
$\zC$ & Gibbs reference vector, $\zC_i=\z_i\exp(-C_i/\gamma)$. \\
$F_\gamma(u)$ & Dual objective associated with the entropically regularized problem. \\
$u=(u_1,u_2)$ & Dual variable split according to the two constraint blocks. \\
$x(u)$ & Primal variable recovered from a dual variable by the primal--dual relation. \\
$\Psi_1,\Psi_2$ & Exact dual block maximization maps. \\
$\Psi=\Psi_1\circ\Psi_2$ & Full dual sweep map on the first dual block. \\
$\|\cdot\|_{V_1},\|\cdot\|_{V_2}$ & Block quotient seminorms induced by $\ker(A^\top)$. \\
$\|\cdot\|_V$ & Maximum of the two block quotient seminorms. \\
$\Delta_k$ & Dual suboptimality gap $F_\gamma^\star-F_\gamma(u^{(k)})$. \\
$\Xmax$ & Uniform $\ell^1$ bound on primal iterates. \\
$\Umax$ & Uniform block-quotient bound on dual iterates. \\
$H_\gamma$ & Uniform bound on $\|\log x_\gamma-\log \z\|_\infty$ at the regularized optimum. \\
$\kappa(A_1,A_2)$ & Decomposition constant converting control of $A^\top u$ into quotient control of dual blocks. \\
$\Vertex,\Edge$ & Vertex and edge sets of the graph used for graph $W_1$. \\
$n,p$ & Number of graph vertices and directed sparse edge entries, respectively. \\
$\length$ & Edge-length/cost matrix on the graph. \\
$\Flows$ & Sparse nonnegative flow cone supported on graph edges. \\
$f,g$ & Duplicated graph-flow variables used in the flow-Sinkhorn splitting. \\
$v,U$ & Vertex and edge dual variables in the graph-flow formulation. \\
$\operatorname{diameter}(\Edge)$ & Maximum shortest-path distance between graph vertices. \\
$\Sigma,\tau$ & Signature matrix and scalar balance parameter used for signed monotonicity and translation equivariance. \\
$D_\phi$ & General Bregman divergence generated by a convex function $\phi$. \\
$\eta_\gamma$ & Generalized Pinsker constant in the Bregman extension. \\

\bottomrule
\end{longtable}
